Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-13 paper/ms.tex
update ms

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CQVs}.  The complete figure
			set, with one image per sector for each of \ncpvsfound\ CQVs
			and CQV candidates
			is available online {\bf For internal collaboration review:
				\url{https://www.dropbox.com/scl/fo/zlj3txot4cvymfb22wewu/h?dl=0&rlkey=3ec5f9o5xewrixzfkhkdenopa}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

Broadly speaking, the most common false positives for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false positive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labeled either ``good'' or ``possible'' CQVs.  The
Fourier analysis was not amenable to calculating a similar
false-positive probability, because it was implemented as a general
variability search; its results are also being used for other types of
variability analyses beyond CQVs.

@@

Table~\ref{tab:thetable} lists the \ncpvsfound\ objects identified by
our search.  The ``high-quality sample'' includes \ngoods\ of these
objects, which demonstrated what we viewed as unambiguous
characteristics of the CQV phenomenon in at least one TESS sector.
The classification of the remaining \nmaybes\ CQV candidates was
ambiguous.  Additional data from TESS or other instruments could help
resolve their classification.  The boolean \texttt{quality} column in
the table divides the two classes.  In the following, we will restrict
our discussion to the high-quality sample.

@@

A small number of objects at first glance seem reminiscent of
eclipsing binaries, such as TIC~193831684 or TIC~5714469.  In these
few cases, we believe that that are unlikely to be eclipsing binaries
due to additional coherent variations in the light curves that are
distinct from any binary phenomenology of which we are aware.

@@

Of our \ncpvsfound\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality, and youth, of that group.  The yield
in Sco-Cen is not surprising, since Sco-Cen contains the majority of
pre-main-sequence stars in the solar neighborhood.  However, given the
$\lesssim$$10\%$ completeness of TESS beyond 100\,pc, there are
probably many more CQVs that remain to be discovered in Sco-Cen.

Of the \ngoodsfieldbanyan\ stars for which BANYAN\,$\Sigma$ did not
find any association, one (TIC~302160226) is a member of $\alpha$~Per
($t\approx 86\pm16$\,Myr;
\citealt{2021A&A...645A..84M,2023AJ....166...14B}).  For the other two
(TIC~58084670 and TIC~141146667), we were not able to confidently
associate either star with any young groups.  However both do seem to
show the CQV signal over multiple TESS sectors, and both are
photometrically elevated relative to the main sequence.  For instance,
both were noted by \citet{2021ApJ...917...23K} as being in the
``diffuse'' population of $<$50\,Myr stars near the Sun.

@@

Most CQVs in our catalog did not show infrared excesses in the W1-W4
bands, which is typical for this class of object
\citep{2017AJ....153..152S}.  Visually inspecting the SEDs of our
\ncpvsfound\ star sample, we labeled two objects as having ``good''
infrared excesses (both W3 and W4 at $>$3$\sigma$ above the
photospheric prediction), and three as ``possible'' infrared excess
(only W3 at $>$3$\sigma$).

@@

Out of the \ngoods\ CQVs, a small fraction were in regions of sky with
many sectors of adjacent TESS data.  Out of these objects, LP 12-502
(TIC 402980664; $d$=21\,pc, $J$=9.4, $T$=11.1) stood out due to the
quality and content of its data.  We describe another interesting
source, TIC~300651846, in Appendix~\ref{app:tic3006}.  In this
section, we first present the LP 12-502 observations
(Section~\ref{subsec:lpobservations}), and then highlight their
implications (Section~\ref{subsec:lplessons}).

@@

One challenge is perhaps that in systems like LP 12-502, the observed
dip depth variations would necessarily drive one toward a picture of a
disintegrating planetesimal swarm, rather than just a single
planetesimal.  The need for more than just one launching body would be
driven by the observation that we see dips appear and disappear at
many different phases, but with roughly the same period.  The observed
changing dip depths in response to flares would be understood through
the planetesimal outflow rates being stalled by changes in the local
magnetic field or coronal plasma density.  The sizes of these
purported planetesimals would need to be $\lesssim 1$\,$R_\oplus$,
based on the non-detections of their transits, analogous to e.g.
K2-22 \citep{2015ApJ...812..112S} or KOI-2700
\citep{2014ApJ...784...40R}.  Typical sizes of bodies in the asteroid
belt for instance span hundreds of meters to a few kilometers.

There are at least three problems with this concept of a
disintegrating planetesimal swarm.  First, the model would predict
that certain orbital phases would produce recurrent dips if observed
over sufficiently long baselines, because the launching planetesimal
would be massive enough to remain in orbit, while stochastically
ejecting material.  An analogous system would be K2-22
\citep{2015ApJ...812..112S}.  For most CQVs (Figure~\ref{fig:evoln}),
the data are in tension with this expectation.  A second issue
is that many of the dips show asymmetries in the wrong direction
relative to the naive expectation of a trailing comet tail.  Invoking
non-exponentially decaying dust distributions as a function of azimuth
might be one way out, but such an idea lacks firm theoretical footing.
Finally, dynamical stability is a concern: in a model in which any new
dip is ``explained'' by invoking a new planetesimal, one might
eventually pack the corotation radius to a degree where instability
over short timescales would be guaranteed.

@@

In our target sample, we found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bonafide CQVs are listed in Table~\ref{tab:thetable}.  This table also
includes an additional \nmaybes\ ambiguous CQVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but two of the \ncpvsfound\
objects based on association memberships; we also derived temperatures
and radii using SED fitting, and then inferred stellar masses by
interpolating against stellar evolutionary models.  We caution that
our selection function was not volume-limited: the TESS 120-second
stellar sample had a heterogeneous selection function which may have
been biased in favor of young stars over field stars.  Previous work
however has shown that $\approx$1-3\% of M dwarfs younger than
$\approx$100\,Myr show the CQV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.

>>>>

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CQVs}.  The complete figure
			set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable}
			is available online {\bf For internal collaboration review:
				\url{https://www.dropbox.com/scl/fo/zlj3txot4cvymfb22wewu/h?dl=0&rlkey=3ec5f9o5xewrixzfkhkdenopa}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

Broadly speaking, the most common false positives for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false positive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labeled either ``good'' or ``possible'' CQVs.  The
Fourier analysis was not amenable to calculating a similar
false-positive rate, because it was implemented as a general
variability search; its results are also being used for analyses
on topics other than CQVs.

@@

Table~\ref{tab:thetable} lists the \nallcands\ objects identified by
our search.  The ``high-quality sample'' includes \ngoods\ of these
objects, which demonstrated what we viewed as unambiguous
characteristics of the CQV phenomenon in at least one TESS sector.
The classification of \nmaybes\ CQV candidates was ambiguous, and the
\ndebunked\ remaining objects were false positives that we discuss below.
The \texttt{quality} column in the table divides the three classes;
additional data from TESS or other instruments could help resolve the
classification of the ambiguous cases.   Of the \ncqvsnodebunked\ CQVs
and candidate CQVs, \nbothdipfourier\ were found using both the
dip-counting and Fourier techniques, \nyesdipnofourier\ were found
using only the dip-counting technique, and \nyesfouriernodip\ were
found using only the Fourier technique.  In the following, we will
restrict our discussion to the high-quality sample, irrespective of
discovery method.

@@

A small number of objects at first glance seem reminiscent of
eclipsing binaries, such as TIC~193831684 or TIC~5714469.  We 
believe these cases are unlikely to be eclipsing binaries
due to the additional coherent peaks and troughs in the light curves,
which are
distinct from any binary phenomena of which we are aware.

@@

Of our \ncqvsnodebunked\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality, and youth, of that group.  The yield
in Sco-Cen is not surprising, since Sco-Cen contains the majority of
pre-main-sequence stars in the solar neighborhood.  However, given the
$\lesssim$$10\%$ completeness of TESS beyond 100\,pc, there are
probably many more CQVs in Sco-Cen that remain to be discovered.

Of the \ngoodsfieldbanyan\ stars for which BANYAN\,$\Sigma$ did not
find any association, one (TIC~302160226) is a member of $\alpha$~Per
($t\approx 86\pm16$\,Myr;
\citealt{2021A&A...645A..84M,2023AJ....166...14B}).  For the other two
(TIC~58084670 and TIC~141146667), we were not able to confidently
associate either star with any young groups.  Both do seem to
show the CQV signal over multiple TESS sectors, and both are
photometrically elevated relative to the main sequence.  
Both were also noted by \citet{2021ApJ...917...23K} as being in the
``diffuse'' population of $<$50\,Myr stars near the Sun.

@@

Most CQVs in our catalog did not show infrared excesses in the W1-W4
bands, which is typical for this class of object
\citep{2017AJ....153..152S}.  Visually inspecting the SEDs of our
\nallcands\ star sample, we labeled two objects as having ``good''
infrared excesses (both W3 and W4 at $>$3$\sigma$ above the
photospheric prediction), and three as ``possible'' infrared excess
(only W3 at $>$3$\sigma$).

@@

A small fraction of our objects were in regions of sky with
many sectors of adjacent TESS data.  Out of this already small sample, LP 12-502
(TIC 402980664; $d$=21\,pc, $J$=9.4, $T$=11.1) stood out due to the
quality and content of its data.  We discuss another interesting
source, TIC~300651846, in Appendix~\ref{app:tic3006}.  In this
section, we first present the LP 12-502 observations
(Section~\ref{subsec:lpobservations}), and then highlight their
implications (Section~\ref{subsec:lplessons}).

@@

In systems like LP 12-502, the observed
variability would probably drive one toward a picture of a
disintegrating planetesimal swarm, rather than just a single
planetesimal.  The need for more than just one launching body 
is because we see dips appear and disappear at
many different phases, but with roughly the same period.  The observed
changing dip depths in response to flares would be understood through
the planetesimal outflow rates being stalled by changes in the local
magnetic field or coronal plasma density.  The sizes of these
purported planetesimals would need to be $\lesssim 1$\,$R_\oplus$,
based on the non-detections of their transits, analogous to e.g.
K2-22 \citep{2015ApJ...812..112S} or KOI-2700
\citep{2014ApJ...784...40R}.  Typical sizes of bodies in the asteroid
belt for instance span hundreds of meters to a few kilometers.

There are at least three problems with this 
disintegrating planetesimal swarm concept.  First, the model would predict
that certain orbital phases would produce recurrent dips if observed
over sufficiently long baselines, because the launching planetesimal
would be massive enough to remain in orbit, while stochastically
ejecting material.  An analogous system would be K2-22
\citep{2015ApJ...812..112S}.  For most CQVs (Figure~\ref{fig:evoln}),
the data are in tension with this expectation.  A second issue
is that many of the dips show asymmetries in the wrong direction
relative to the naive expectation of a trailing comet tail.  Invoking
non-exponentially decaying dust distributions as a function of azimuth
might be one way out, but such an idea lacks firm theoretical footing.
Finally, dynamical stability is a concern: in a model in which any new
dip is ``explained'' by invoking a new planetesimal, one might
eventually pack the corotation radius to a degree where instability
over short timescales would be guaranteed.

@@

In our target sample, we found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bonafide CQVs are listed in Table~\ref{tab:thetable}.  This table also
includes an additional \nmaybes\ ambiguous CQVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but two of the \nallcands\
objects based on association memberships; we also derived temperatures
and radii using SED fitting, and then inferred stellar masses by
interpolating against stellar evolutionary models.  We caution that
our selection function was not volume-limited: the TESS 120-second
stellar sample had a heterogeneous selection function which may have
been biased in favor of young stars over field stars.  Previous work
however has shown that $\approx$1-3\% of M dwarfs younger than
$\approx$100\,Myr show the CQV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.
